On The Impact of Atoms of Confusion in JavaScript Code

### 2021-04-25 4_Result.tex
Updates from Overleaf

\subsubsection*{Hypothesis Testing}
We use the non-parametric \emph{Mann-Whitney test} to
investigate the null hypothesis that developers 
spend the same amount of time to correctly
predict the outcome of a code snippet, regardless
of evaluating the confusing or  clean version
of the code. The results of Table~\ref{tab:difference-time-taken}
suggest that we should refute the null hypotheses for
five atom candidates (Comma Operator, Logic
as Control Flow, Implicit Predicate, Arithmetic
as Logic, and Pre-Increment). This result holds even after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%. 
For the first three, the analysis suggests that the participants
need more time to predict the outcome of
the code snippet in the confusing code version. Interesting,
for the atom candidates Arithmetic as Logic and
Pre-Increment, the participants take less time
to predict the outcome of the code snippets in
the confusing version.
We also computed the effect size using the Cliff's Delta method.
We found a large effect for the Comma Operator atom
candidate; and a moderate effect for the atom candidates
Logic as Control Flow, Implicit Predicate, Arithmetic
as Logic, and Pre-Increment.

@@

% latex table generated in R 4.0.3 by xtable 1.8-4 package
% Tue Apr 13 09:34:35 2021
\begin{table}[ht]
\centering
\caption{Summary of atoms occurrences on our dataset}
\setlength\tabcolsep{2pt} % default value: 6pt
\label{tab:occurrences-summary}
\begin{tabular}{lrcc}
  \hline
Atom & Projects & Occurrences/KLOC & Category \\ 
  \hline
Implicit Predicate & 100.00 & 19.89 &  highly used \\ 
  Ternary Operator & 100.00 & 10.16 &  highly used \\ 
  Omitted Curly Braces & 91.67 & 6.61 &  highly used \\ 
  Post Increment & 90.28 & 5.62 &  highly used \\ 
  Pre Increment & 83.33 & 1.04 & commonly used \\ 
  Assignment as Value & 81.94 & 1.02 & commonly used \\ 
  Logic as Control Flow & 56.94 & 0.95 & commonly used \\ 
  Arithmetic as Logic & 23.61 & 0.02 & little used \\ 
  Comma Operator & 15.28 & 0.03 & little used \\ 
   \hline
\end{tabular}
\end{table}

Considering all JavaScript projects in our dataset, we found 
a total of {\color{red}\num{363810}} atoms of confusion, though three atoms are responsible for 87.97\% of this total: Ternary Operator, Omitted Curly Braces and the Post Increment. The remaining atoms
account for \num{24558} (12.03\%) of the total number of occurrences, and the Comma Operator and Arithmetic as Logic are the atoms that arise less frequently (0.001\% in total with 232 and 165 occurrences, respectively).

>>>>

\subsubsection*{Hypothesis Testing}
We use the non-parametric \emph{Mann-Whitney U test}, also known as Wilcoxon rank-sum test, to
investigate the null hypothesis that developers 
spend the same amount of time to correctly
predict the outcome of a code snippet, regardless
of evaluating the confusing or  clean version
of the code. The results of Table~\ref{tab:difference-time-taken}
suggest that we should refute the null hypotheses for
five atom candidates (Comma Operator, Logic
as Control Flow, Implicit Predicate, Arithmetic
as Logic, and Pre-Increment). This result holds even after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%. 
For the first three, the analysis suggests that the participants
need more time to predict the outcome of
the code snippet in the confusing code version. Interesting,
for the atom candidates Arithmetic as Logic and
Pre-Increment, the participants take less time
to predict the outcome of the code snippets in
the confusing version.
We also computed the effect size using the Cliff's Delta method.
We found a large effect for the Comma Operator atom
candidate; and a moderate effect for the atom candidates
Logic as Control Flow, Implicit Predicate, Arithmetic
as Logic, and Pre-Increment.

@@

% latex table generated in R 4.0.3 by xtable 1.8-4 package
% Tue Apr 13 09:34:35 2021
\begin{table}[ht]
\centering
\caption{Summary of atoms occurrences on our dataset}
\setlength\tabcolsep{2pt} % default value: 6pt
\label{tab:occurrences-summary}
\begin{tabular}{lrcc}
  \hline
Atom & Projects & Occurrences/KLOC & Category \\ 
  \hline
Implicit Predicate & 100\% & 19.89 &  highly used \\ 
  Ternary Operator & 100\% & 10.16 &  highly used \\ 
  Omitted Curly Braces & 91.67\% & 6.61 &  highly used \\ 
  Post Increment & 90.28\% & 5.62 &  highly used \\ 
  Pre Increment & 83.33\% & 1.04 & commonly used \\ 
  Assignment as Value & 81.94\% & 1.02 & commonly used \\ 
  Logic as Control Flow & 56.94\% & 0.95 & commonly used \\ 
  Automatic Semicolon Insertion & 33.33\% & 0.13 & commonly used \\ 
  Arithmetic as Logic & 23.61\% & 0.02 & little used \\ 
  Comma Operator & 15.28\% & 0.03 & little used \\ 
   \hline
\end{tabular}
\end{table}

Considering all JavaScript projects in our dataset, we found 
a total of {\color{red}\num{364873}} atoms of confusion, though three atoms are responsible for 87.97\% of this total: Ternary Operator, Omitted Curly Braces and the Post Increment. The remaining atoms
account for \num{24558} (12.03\%) of the total number of occurrences, and the Comma Operator and Arithmetic as Logic are the atoms that arise less frequently (0.001\% in total with 232 and 165 occurrences, respectively).

### 2021-04-25 5_Threats.tex
Updates from Overleaf

%% Since our research was conducted online with unknown participants, and thus we had no way of confirming their
%% levels of education and experience. Also,  we did not have a way to prevent respondents from cheating,
%% such as running the code on an interpreter, or consulting other people. This presents a threat to the validity to the
%% conclusions of our work, even though we leverage some means to try to mitigate this threat (as discussed
%% in Section~\ref{method}).

%% Another threat to our results lies in the fact that some constructs
%% might not be frequent, or present at all, in JavaScript code bases.
%% The atom candidates Arithmetic as Logic and
%% the Comma Operator were rarely found in the
%% open source JavaScript code we mined in our
%% third study.

%% A potential problem with our method is that
%% there little incentive for subjects to think
%% thoroughly about the questions. We observed a lack of
%% engagement when we ran the survey with undergraduate students during
%% a pilot study. Although our final subjects were voluntarily
%% partaking in the survey, we could not be sure that, after some
%% time taking the survey, respondents would become tired and stop thinking
%% clearly about the code.

>>>>

Conclusion validity is connected with how well it is possible to establish relationships between treatments and outcomes. Threats to conclusion validity often come from inappropriate use of statistics. In the context of our survey study, we apply different statistical tests appropriate for the cases where data was categorical (correctness, Chi-square test of independence) and continuous (time, Mann-Whitney U test). Furthermore, besides reporting p-values, we also report effect size measures appropriate for each scenario (odds-ratio and Cliff's delta) and apply a p-value correction technique to avoid the multiple testing problem. Finally, it could be argued that the size of the samples is insufficient to make conclusions for some of the atoms, a common problem in empirical software engineering. We estimate the sample size for each one of the atom candidates, considering that each one had a different number of samples (Table~\ref{tab:difference-correctness}). To that end, we use the $\phi$ measure of effect size for each atom (based on the Chi-square statistic), the standard $\alpha$ coefficient of 0.05, and set the expected statistical power to 0.8, as usually employed in the literature~\cite{Ellis:2010:EGE}. As a result of this analysis, we discover that for all but one of the atoms where we had a statistically significant result for correctness, Implicit Predicate, the sample size is sufficient. For the ones where we did not find statistically significant results for correctness, in fact the sample sizes we have analyzed are insufficient for such small effect sizes. This indicates that further studies on these atoms are required.

%One threat to validity we have not properly addressed is lack of power, a common problem in Software Engineering research.

Construct validity is connected with how well the selected measures actually represent the concept of interest. We use correctness as a proxy for program comprehension. Furthermore, we use predicted outcomes of small programs as a measure of correctness. As discussed elsewhere~\cite{Oliveira:2020:ECR}, this approach is a test of the developers' ability to trace programs. Although this is a common approach in studies about atoms of confusion~\cite{TheEyesDoNotLie,Langhout:2021:ACJ,DBLP:conf/sigsoft/GopsteinIYDZYC17}, other measures of correctness could have been employed, potentially yielding different results. As a complement to correctness in the survey, we have measured the time it took for the participants to finish. Furthermore, we have  asked the interviewees about their preferences when comparing confusing and clean versions of the programs.

%A potential problem with our method is that there little incentive for subjects to think thoroughly about the questions. We observed a lack of engagement when we ran the survey with undergraduate students during a pilot study. Although our final subjects were voluntarily partaking in the survey, we could not be sure that, after some time taking the survey, respondents would become tired and stop thinking clearly about the code.

Internal validity is concerned with how well the study isolates the variables of interest and accounts for confounding factors. Since the survey was conducted online with unknown participants, we have no way of confirming their levels of education and experience. We mitigate this threat by, in the data analysis, explicitly accounting for programmer experience and using an experimental design that allows us to isolate the impact of the treatment from factors such as experience and formal education level. Also, we did not have a way to prevent respondents from cheating, such as running the code on an interpreter, or consulting other people. We partially mitigate this threat by presenting images with source code, instead of text. This creates an obstacle for participants to run the code while taking the survey.

External validity is linked to whether it is possible to extrapolate the results of the study. Although our results suggest that the selected idioms and code constructs may lead to confusion for small code snippets, it is not clear whether that result extrapolates to other scenarios. Even though it is likely that in larger code bases the confusion induced by these constructs may be even greater, the existence of additional context may mitigate this effect. Another possible threat to the generalizability of our results, in particular for the survey and interviews, lies in the fact that the analyzed atoms may rarely occur in real JavaScript code bases. To mitigate that threat, we have analyzed 72 popular open source JavaScript repositories and found out that most of them are fairly frequent, occurring at least once per 1,000 lines of code. Only Arithmetic as Logic and Comma Operator occur less frequently than once per 10,000 lines of code.

### 2021-04-25 main.tex
Updates from Overleaf

\begin{abstract}
Evolving legacy code is a challenging task, particularly when the code has been poorly written---which imposes a significant cognitive load on developers when trying to familiarize themselves with a legacy codebase. Previous research has investigated the impact of small code patterns that contribute to misunderstanding the source code written in C and C++. However, little is known about the influence of these small code patterns (hereafter atoms of confusion) on understanding dynamic languages (such as JavaScript). In this work, we present the results of a mixed-method research: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort.
Our goal is to investigate the impact of atoms of confusion on understanding code written in JavaScript --- a dynamically typed language whose popularity is growing in the most different application domains (from web programming to backend and IOT development). From our survey and interviews, we confirm that atoms of confusion lead to code that is hard to understand: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in 6 out of 9 atom candidates. Also, the results of our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates
appear in more than 60\% of the projects---including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively.
Our investigation also reveals additional JavaScript constructs and idioms that might hinder developers from understanding JavaScript code. For instance, even experienced JavaScript developers agree that it is hard to understand code that uses callback functions, even though we found widespread use of callbacks in our projects dataset. Altogether, our findings might help developers better understand the implications of atoms of confusion on understanding JavaScript code and contribute to the design of new program transformation tools that aim to simplify JavaScript code.
\end{abstract}

>>>>

\begin{abstract}
Evolving legacy code is a challenging task, particularly when the code has been poorly written, increasing effort and imposing a significant cognitive load on developers. Previous research has investigated the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in C and C++. However, little is known about the influence of these atoms on understanding dynamic languages (such as JavaScript). In this work, we investigate the impact of atoms of confusion on understanding code written in JavaScript in terms of a mixed-method research: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort. From our survey and interviews, we confirm that atoms of confusion lead to code that is hard to understand: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in 6 out of 9 atom candidates. Our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects, including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively. Altogether, our findings might help developers better understand the implications of atoms of confusion on understanding JavaScript code and contribute to the design of program transformation tools to simplify JavaScript code.

% Our investigation also reveals additional JavaScript constructs that might hinder even experienced developers from understanding JavaScript code, such as the use of callbacks functions. Altogether...

%Evolving legacy code is a challenging task, particularly when the code has been poorly written---which imposes a significant cognitive load on developers when trying to familiarize themselves with a legacy codebase. Previous research has investigated the impact of small code patterns that contribute to misunderstanding the source code written in C and C++. However, little is known about the influence of these small code patterns (hereafter atoms of confusion) on understanding dynamic languages (such as JavaScript). In this work, we present the results of a mixed-method research: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort. Our goal is to investigate the impact of atoms of confusion on understanding code written in JavaScript --- a dynamically typed language whose popularity is growing in the most different application domains (from web programming to backend and IOT development). From our survey and interviews, we confirm that atoms of confusion lead to code that is hard to understand: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in 6 out of 9 atom candidates. Also, the results of our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects---including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively. Our investigation also reveals additional JavaScript constructs and idioms that might hinder developers from understanding JavaScript code. For instance, even experienced JavaScript developers agree that it is hard to understand code that uses callback functions, even though we found widespread use of callbacks in our projects dataset. Altogether, our findings might help developers better understand the implications of atoms of confusion on understanding JavaScript code and contribute to the design of new program transformation tools that aim to simplify JavaScript code.
\end{abstract}
